Effectiveness of community-based marine reserves in small-scale fisheries

### 2018-09-18 docs/paper/Villasenor-Derbez.tex
Updated manuscript, almost ready for submission

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,2,*}\), Eréndira
Aceves-Bueno\(^{1}\), Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\),
Arturo Hernández-Velasco\(^{2}\), Jorge Torre\(^{2}\), Fiorenza
Micheli\&\textsuperscript{\{3\}}\&}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas,
Mexico\newline \(^{3}\)Hopkins Marine Station and Center for Ocean
Solutions, Stanford University, Pacific Grove, California}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

\title[Community-based marine reserves: 2018-09-13]{Effectiveness of community-based marine reserves in small-scale
fisheries}

@@

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Artisanal fisheries face great
challenges since they are difficult to monitor, enforce, and manage.
Combining territorial user rights for fisheries (TURF) with no-take
marine reserves to create TURF-reserves is believed to improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last 4
years, 45 TURF reserves have been implemented in Mexico, but their
effectiveness has not been formally evaluated. We combine causal
inference techniques and a social-ecological systems framework to
provide a holistic evaluation of community-based TURF reserves in three
coastal communities in Mexico. We find that while reserves have not yet
achieved their stated goal of increasing lobster densities, they
continue to receive significant support from the fishing communities. A
lack of ecological and socioeconomic effects or reserves likely results
from a combination of factors. First, the lobster fisheries are already
well managed, and it is unlikely that reserves might have a detectable
effect. Second, reserves are not large enough to protect lobster's home
ranges. Third, some of these reserves might be too young for the effects
to show. However, these reserves have shaped small-scale fishers' way of
thinking about marine management, which can provide a foundation for
establishing additional, larger marine reserves needed to effectively
conserve mobile species.

\medskip
\tiny
 \keyFont{ \section{Keywords:} Marine Protected Areas, Marine Conservation, Small-Scale Fisheries,
Citizen Science, TURF-reserves, Social-Ecological Systems}

@@

TURF-reserves can be created as community-based marine reserves,
voluntarily established and enforced by local communities. This
bottom-up approach increases compliance and self-enforcement
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
Community-based spatial closures occur in different contexts, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, without legal recognition
no-take regulations are difficult to enforce and fishers rely on the
exclusive access granted by the TURF. In an effort to bridge this
normative gap, Civil Society Organizations (CSOs) have served as a link
between fishers and government, and set out to create a legal framework
that solve this governance issue. In Mexico, a new norm was created in
2014 allowing fishers to request the legal recognition of
community-based reserves as ``Fish Refuge'' (\emph{Zona de Refugio
Pesquero} ;\citet{nom}). Fish refuges can be implemented as temporal or
partial reserves, which can protect one, some, or all resources within
their boundaries. Since 2014, 45 of community-based marine reserves
along the Pacific, Gulf of California, and Mexican Caribbean coastlines
have gained legal recognition. However, their effectiveness has not yet
been formally evaluated and reported in the scientific literature.

@@

\section{Materials and Methods}\label{materials-and-methods}

@@

The other two TURF-reserves are located in Maria Elena and Punta
Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
Isla Nativdad, which hosts a well established fishing community, Maria
Elena is a fishing camp --visited intermittently during the fishing
season-- belonging to the Cozumel fishing cooperative (\emph{SCPP
Cozumel}); Punta Herrero is home to the \emph{SCPP José María Azcorra}
cooperative, and similar to Isla Natividad hosts a local community.
Their main fishery is the Caribbean spiny lobster (\emph{Panulirus
argus}), but they also target finfish in the off-season. Maria Elena and
Punta Herrero established eight marine reserves in 2012, and four marine
reserves in 2013, respectively. All these reserves are legally
recognized as Fishing Refugia since their creation.

These communities are representative of their region in terms of
ecology, socioeconomic, and governance aspects. Isla Natividad, for
example, is part of a greater group of fishing cooperatives belonging to
a Federation of Fishing Cooperatives. This group has been identified as
a cohesive group that often cooperates to better manage their resources
\citep{mccay_2017-1m,mccay_2014-CN,acevesbueno_2017}. Likewise, Maria
Elena and Punta Herrero are representative of fishing cooperatives in
the Mexican Caribbean, which are also part of a regional Federation.
Together, these three communities provide an accurate representation of
other fishing communities in each of their regions. While each region
has more communities that have established community-based
TURF-reserves, available data would not allow us to perform the in-depth
analysis that we undertake. Yet, given the similarities among
communities and the socioeconomic and governance setting under which
they opperate, it is safe to cautiously generalize our results to other
communities in Mexico and other regions around the world.

@@

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control areas, carried out by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species along replicate transects
(30x 2 m each) at depths 5-20 m in the reserves and control sites
\citep{fulton_2018}. Size structures are also collected during fish
surveys. We define control sites as regions with habitat characteristics
similar to the corresponding reserves, and that presumably had a similar
probability of being selected as reserves during the design phase. We
focus our evaluation on sites where data are available for reserve and
control sites, before and after the implementation of the reserve. This
provides us with a Before-After-Control-Impact (\emph{i.e.} BACI)
sampling design that allows us to capture and control for temporal and
spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI designs and
causal inference techniques have proven effective to evaluate marine
reserves, as they allow us to causally attribute observed changes to the
intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites
were surveyed annually, and at least once before implementation of the
reserves. Table \ref{table:com_sum} shows a summary of the TURF-reserves
included in this study.

@@

\caption{\label{tab:unnamed-chunk-2}\label{table:com_sum} Summary of commuity--based marine reserves by community.}
\centering
\begin{tabular}[t]{l|r|r|r|r}
\hline
Community & TURF area ($\mathrm{km}^{-2}$) & Reserve area ($\mathrm{km}^{-2}$) & Percent as reserves & Year of implementation\\
\hline
Isla Natividad & 889.5 & 1.53 & 0.1720067 & 2006\\
\hline
Maria Elena & 353.1 & 0.10 & 0.0283206 & 2012\\
\hline
Punta Herrero & 299.7 & 0.43 & 0.1434768 & 2013\\
\hline
\end{tabular}
\end{table}

Socioeconomic data come from landing receipts reported to the National
Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
(Kg) and revenues (MXP) for cooperatives with and without marine
reserves(\textbf{Fig S1}). Cooperatives incorporated in this analysis
belong to larger regional-level Cooperative Federations, and are exposed
to the same markets and institutional frameworks, making them plausible
controls \citep{mccay_2017-1m,ayer_2018}. Landings and revenues were
aggregated at the cooperative-year level, and revenues were adjusted to
represent 2014 values by the Consumer Price Index for Mexico
\citep{oecd_2017-VV} as:

@@

Data for the operationalization of the social-ecological system were
collected at the community--level from official documents used in the
creation and designation of the marine reserves
\citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on
the authors' experience and knowledge of the communities. These include
information on the resource system, the resource units, actors, and the
governance system itself (\textbf{S1 Table}).

@@

We use the social-ecologycal system (SES) framework to evaluate each
community as a means of providing an explanation to the biological and
socioeconomic results. The use the SES framework standardizes our
analysis and allows us to communicate our results in a common language
across fields. We based our variable selection primarily on
\citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed
Mexican fishing cooperatives using this framework. \textbf{S1 Table}
shows the selected variables, the definition and selected indicators and
scoring system.

@@

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and forth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). Likewise, no changes were detected in fish biomass or
invertebrate and fish densities (\ref{fig:indicators}B-D), where effect
sizes oscillated around zero without clear trends. Full tables with
model coefficients are presented in the supplementary materials
(\textbf{S2 Table}, \textbf{S3 Table}, \textbf{S4 Table}).

@@

Lobster landings and revenue were only available for Isla Natividad and
Maria Elena (Fig \ref{fig:lobsters}). For all years before
implementation, the effect sizes are close to zero, indicating that the
control and treatment sites have similar pre-treatment trends,
suggesting that these are plausible controls. However, effect sizes do
not change after the implementation of the reserve. Interestingly, the
negative effect observed for Isla Natividad on year 5 correspond to the
2011 hypoxia events. The only positive change observed in lobster
landings is for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
post-implementation data for Maria Elena do not show a significant
effect of the reserve. Isla Natividad shows higher revenues after the
implementation of the reserve, as compared to the control communities.
However, these changes are not significant and are associated to
increased variation. All regression coefficients for each community and
indicator are presented in \textbf{S5 Table}.

@@

Although we have little information on the social dimension of these
fisheries, we can use the social-ecological systems framework
(\textbf{S1 Table}) to analyze the performance of each governance system
(\textbf{S6 Table}). Our analysis shows that all communities analyzed
share similarities in their Governance system which is based on
cooperatives (GS5.2.3.2), with strong rules in use that include
Operational rules (GS6.2), Collective-choice rules (GS6.3),
Constitutional rules (GS6.3), and even Territorial use communal rights
(GS6.1.4.3). However, we identified important differences in terms of
the actors, resource systems, and resource units. Although all
communities show a high level of leadership (A5), the level of trust
(A6.1) is lower in Punta Herrero. In general, the presence and success
of conservation initiatives depends on the incentives of local
communities to maintain a healthy status of the resources they depend
upon \citep{jupiter_2017}. The enabling conditions for conservation seem
to be strongly present in all communities. Due to the clarity of access
rights and isolation, the benefits of conservation directly benefit the
members of the fishing cooperative. These conditions have favored the
development of efficient community-based enforcement systems.

@@

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:unnamed-chunk-6}\label{fig:lobsters}Effect sizes for
lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
circles) and Maria Elena (ME; blue triangles). Points indicate the
effect size and standard errors. Years have been centered to year of
implementation.}
\end{figure}

@@

Some works evaluate marine reserves by performing inside-outside
\citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or
before-after comparisons \citep{betti_2017-lq}. The first approach does
not address temporal variability, and the second can not distinguish
between the temporal trends in a reserve and the entire system
\citep{depalma_2018}. Our approach to evaluate the temporal and spatial
changes provides a more robust measure of reserve effectiveness. For
example, we capture previously described patterns like the rapid
increase observed for changes in lobster densities for Isla Natividad on
the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A) occur a
year after the hypoxia events described by \citet{micheli_2012-EU},
which caused mass mortality of sedentary organisms such as abalone and
sea urchins, but not lobster and finfish. Yet, our empirical approach
assumes control sites are a plausible counterfactual for treated sites.
This implies that treated sites would have followed the same trend as
control sites, had the reserves not been implemented. Nonetheless,
overall trends for each site don't show any significant increases,
supporting our findings of lack of change in the indicators used
(\textbf{S2 Figure}, \textbf{S3 Figure}, \textbf{S4 Figure}, \textbf{S5
Figure}, \textbf{S6 Figure}).

@@

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to poor ecological coherence in reserve dsign (\emph{sensu}
\citet{rees_2018}. Previous research has shown that reserves in Isla
Natividad yield fishery benefits for the abalone fishery
\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
perhaps the reserves provide enough protection to these sedentary
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, may favor
implementation of reserves that pose low fishing costs due to their
small size or location. Our analysis of economic data supports this, as
neither landings nor revenues showed the expected short-term costs
associated to the first years of reserve implementation
\citep{ovando_2016-Wg}.

@@

While reserves have failed to provide fishery benefits up to now, there
are a number of additional ecological, fisheries, and social benefits.
Marine reserves provide protection to a wider range of species and
vulnerable habitat, like coral reefs. These sites can serve as an
insurance against uncertainty and errors in fisheries management, as
well as environmental shocks
\citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (\textbf{, Aalto et
al, in press Am Nat}). Self-regulation of fishing effort (\emph{i.e.}
reduction in harvest) can serve as a way to compensate for future
declines associated to environmental variation \citep{finkbeiner_2018}.
Embarking in a marine conservation project can bring the community
together, which promotes social cohesion and builds social capital.
Furthermore, showing commitment to marine conservation and sustainable
fishing practices allows fishers to have greater bargaining power and
leverage over fisheries management \citep{prezramrez_2012}.

@@

\textbf{Tell me about your funding}

JCVD: CONACyT (Beca de Posgrados en el extranjero, CVU 669403) and the
Latin American Fisheries Fellowship.

AS, AHV, SF and JT Walton Family Foundation, Summit Foundation, and Oak
Foundation.

@@

\section*{Supplemental Data}

\href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material}
should be uploaded separately on submission, if there are Supplementary
Figures, please include the caption in the same file as the figure.
LaTeX Supplementary Material templates can be found in the Frontiers
LaTeX folder

\paragraph*{S1 Figure}
\label{S1_Figure}

Map of control and treated sites in A and control and treated landings
in B

\paragraph*{S2 Figure}
\label{S2_Figure}

Time series of biological indicators for IN

\paragraph*{S3 Figure}
\label{S3_Figure}

Time series of biological indicators for ME

\paragraph*{S4 Figure}
\label{S4_Figure}

Time series of biological indicators for PH

\paragraph*{S5 Figure}
\label{S5_Figure}

Time series of economic indicators for ME

\paragraph*{S6 Figure}
\label{S6_Figure}

Time series of economic indicators for PH

\paragraph*{S1 Table}
\label{S1_Table}

Coefficient estimates for biological indicators in Isla Natividad

\paragraph*{S2 Table}
\label{S2_Table}

Coefficient estimates for biological indicators in Maria Elena

\paragraph*{S3 Table}
\label{S3_Table}

Coefficient estimates for biological indicators in Punta Herrero

\paragraph*{S4 Table}
\label{S4_Table}

Coefficient estimates for economic indicators

>>>>

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,2,*}\), Eréndira
Aceves-Bueno\(^{1}\), Álvin Suarez\(^{2}\), Stuart Fulton\(^{2}\),
Arturo Hernández-Velasco\(^{2}\), Jorge Torre\(^{2}\), Fiorenza
Micheli\(^{3}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Sonora,
Mexico\newline \(^{3}\)Hopkins Marine Station and Center for Ocean
Solutions, Stanford University, Pacific Grove, CA, USA}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

\title[Community-based marine reserves]{Effectiveness of community-based marine reserves in small-scale
fisheries}

@@

Coastal marine ecosystems provide livelihoods for small-scale fishers
and coastal communities around the world. Artisanal fisheries face great
challenges since they are difficult to monitor, enforce, and manage.
Combining territorial user rights for fisheries (TURF) with no-take
marine reserves to create TURF-reserves is believed to improve the
performance of small-scale fisheries by buffering fisheries from
environmental variability and management errors, while ensuring that
fishers reap the benefits of conservation investments. In the last six
years, and following a 2012 regulation, 18 TURF-reserves have been
implemented in Mexico; their effectiveness has not been formally
evaluated. We combine causal inference techniques and a
social-ecological systems framework to provide a holistic evaluation of
community-based TURF reserves in three coastal communities in Mexico. We
find that while reserves have not yet achieved their stated goal of
increasing lobster densities, they continue to receive significant
support from the fishing communities. A lack of ecological and
socioeconomic effects likely results from a combination of factors.
First, the lobster fisheries are already well managed, and it is
unlikely that reserves might have a detectable effect. Second, some of
the reserves are not large enough to protect lobsters' home ranges.
Third, some of these reserves might be too young for the effects to
show. However, these reserves have shaped small-scale fishers' way of
thinking about marine conservation, which can provide a foundation for
establishing additional, larger marine reserves needed to effectively
conserve mobile species.

\medskip
\tiny
 \keyFont{ \section{Keywords:} TURF-reserves, Causal Inference, Social-Ecological Systems, Marine
Protected Areas, Marine Conservation, Small-Scale Fisheries}

@@

TURF-reserves can be created as community-based marine reserves,
voluntarily established and enforced by local communities. This
bottom-up approach increases compliance and self-enforcement
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}.
Community-based spatial closures occur in different contexts, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, without legal recognition
no-take regulations are difficult to enforce and fishers rely on the
exclusive access granted by the TURF. In an effort to bridge this
normative gap, Civil Society Organizations (CSOs) have served as a link
between fishers and government, and set out to create a legal framework
that solve this governance issue. In Mexico, a new norm was created in
2014 allowing fishers to request the legal recognition of
community-based reserves as ``Fish Refuge'' (\emph{Zona de Refugio
Pesquero}; \citet{nom}). Fish refuges can be implemented as temporal or
partial reserves, which can protect one, some, or all resources within
their boundaries. Since 2012, 45 of Fish Refuges have been created along
the Pacific, Gulf of California, and Mexican Caribbean coastlines, with
18 of them implemented as TURF-reserves. However, their effectiveness
has not yet been formally evaluated and reported in the scientific
literature.

@@

\section{Methods}\label{methods}

@@

The other two TURF-reserves are located in Maria Elena and Punta
Herrero, in the Yucatan Peninsula (Fig \ref{fig:map}C). In contrast with
Isla Nativdad, which hosts a well established fishing community, Maria
Elena is a fishing camp --visited intermittently during the fishing
season-- belonging to the Cozumel fishing cooperative (\emph{SCPP
Cozumel}); Punta Herrero is home to the \emph{SCPP José María Azcorra}
cooperative, and similar to Isla Natividad hosts a local community.
Their main fishery is the Caribbean spiny lobster (\emph{Panulirus
argus}), but they also target finfish in the off-season. Maria Elena and
Punta Herrero established eight marine reserves in 2012, and four marine
reserves in 2013, respectively. All these reserves have been legally
recognized as Fishing Refuges since their creation
\citep{dof_website_2012,dof_website_2013}.

These communities are representative of their region in terms of
ecology, socioeconomic, and governance aspects. Isla Natividad, for
example, is part of a greater group of fishing cooperatives belonging to
a Federation of Fishing Cooperatives. This group has been identified as
a cohesive group that often cooperates to better manage their resources
\citep{mccay_2017-1m,mccay_2014-CN,acevesbueno_2017}. Likewise, Maria
Elena and Punta Herrero are representative of fishing cooperatives in
the Mexican Caribbean, which are also part of a regional Federation.
Together, these three communities provide an accurate representation of
other fishing communities in each of their regions. While each region
has additional communities that have established community-based
TURF-reserves, available data would not allow us to perform the in-depth
analysis that we undertake. Yet, given the similarities among
communities and the socioeconomic and governance setting under which
they opperate, it is safe to cautiously generalize our results to other
communities in Mexico and other regions around the world.

@@

We use three main sources of information to evaluate these reserves
across the ecological, socioeconomic, and governance dimensions.
Ecological data come from the annual ecological monitoring of reserve
and control areas, carried out by members from each community and
personnel from the Mexican CSO \emph{Comunidad y Biodiversidad}
(\href{www.cobi.org.mx}{COBI}). Trained divers record richness and
abundances of fish and invertebrate species along replicate transects
(30x 2 m each) at depths 5-20 m in the reserves and control sites
\citep{fulton_2018,fulton_2019}. Size structures are also collected
during fish surveys. We define control sites as regions with habitat
characteristics similar to the corresponding reserves, and that
presumably had a similar probability of being selected as reserves
during the design phase. We focus our evaluation on sites where data are
available for reserve and control sites, before and after the
implementation of the reserve. This provides us with a
Before-After-Control-Impact (\emph{i.e.} BACI) sampling design that
allows us to capture and control for temporal and spatial dynamics
\citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference
techniques have proven effective to evaluate marine reserves, as they
allow us to causally attribute observed changes to the intervention
\citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed
annually, and at least once before implementation of the reserves. Table
\ref{table:com_sum} shows a summary of the TURF-reserves included in
this study.

@@

\caption{\label{tab:unnamed-chunk-2}\label{table:com_sum} Summary of analyzed TURF-reserves by community.}
\centering
\begin{tabular}[t]{l|r|r|l|r}
\hline
Community & TURF area ($\mathrm{km}^{-2}$) & Reserve area ($\mathrm{km}^{-2}$) & Percent as reserves & Year of implementation\\
\hline
Isla Natividad & 889.5 & 1.53 & 0.17\% & 2006\\
\hline
Maria Elena & 353.1 & 0.10 & 0.03\% & 2012\\
\hline
Punta Herrero & 299.7 & 0.43 & 0.14\% & 2013\\
\hline
\end{tabular}
\end{table}

Socioeconomic data come from landing receipts reported to the National
Commission for Aquaculture and Fisheries (\emph{Comisión Nacional de
Acuacultura y Pesca}; CONAPESCA). Data contain monthly lobster landings
(Kg) and revenues (MXP) for cooperatives with and without marine
reserves. Cooperatives incorporated in this analysis belong to larger
regional-level Cooperative Federations, and are exposed to the same
markets and institutional frameworks, making them plausible controls
\citep{mccay_2017-1m,ayer_2018}. Landings and revenues were aggregated
at the cooperative-year level, and revenues were adjusted to represent
2014 values by the Consumer Price Index for Mexico \citep{oecd_2017-VV}
as:

@@

Data for the operationalization of the social-ecological system were
collected at the community--level from official documents used in the
creation and designation of the marine reserves
\citep{dof_website_2012,dof_website_2013,dof_website_2018} and based on
the authors' experience and knowledge of the communities. These include
information on the resource system, the resource units, actors, and the
governance system itself (Table \ref{tab:ses}).

@@

We use the social-ecologycal system (SES) framework to evaluate each
community as a means of providing an explanation to the biological and
socioeconomic results. The use the SES framework standardizes our
analysis and allows us to communicate our results in a common language
across fields. We based our variable selection primarily on
\citet{leslie_2015-na,basurto_2013-oq}, who have previously analyzed
Mexican fishing cooperatives using this framework. Table \ref{tab:ses}
shows the selected variables, the definition and selected indicators,
and scoring system.

@@

Indicators showed ambiguous responses through time for each reserve.
Figure \ref{fig:indicators}A shows positive effect sizes for lobster
densities in Isla Natividad and Punta Herrero during the first years,
but the effect is eroded through time. In the case of Maria Elena,
positive changes were observed in the third and forth year. These
effects are in the order of 0.2 extra organisms \(\mathrm{m}^{-2}\) for
Isla Natividad and Punta Herrero, and 0.01 organisms \(\mathrm{m}^{-2}\)
for Maria Elena, but are not significantly different from zero
(\(p > 0.05\)). Likewise, no changes were detected in fish biomass or
invertebrate and fish densities (\ref{fig:indicators}B-D), where effect
sizes oscillated around zero without clear trends. Full tables with
model coefficients are presented in the supplementary materials (S1
Table, S2 Table, S3 Table).

@@

Lobster landings and revenue were only available for Isla Natividad and
Maria Elena (Fig \ref{fig:lobsters}). For all years before
implementation, the effect sizes are close to zero, indicating that the
control and treatment sites have similar pre-treatment trends,
suggesting that these are plausible controls. However, effect sizes do
not change after the implementation of the reserve. Interestingly, the
negative effect observed for Isla Natividad on year 5 correspond to the
2011 hypoxia events. The only positive change observed in lobster
landings is for Isla Natividad in 2014 (\(p < 0.1\)). The three years of
post-implementation data for Maria Elena do not show a significant
effect of the reserve. Isla Natividad shows higher revenues after the
implementation of the reserve, as compared to the control communities.
However, these changes are not significant and are associated to
increased variation. Full tables with model coefficients are presented
in the supplementary materials (S4 Table, S5 Table.

@@

Although we have little information on the social dimension of these
fisheries, we can use the social-ecological systems framework to analyze
the performance of each governance system (Table \ref{tab:ses}). Our
analysis shows that all communities analyzed share similarities in their
Governance system which is based on cooperatives (GS5.2.3.2), with
strong rules in use that include Operational rules (GS6.2),
Collective-choice rules (GS6.3), Constitutional rules (GS6.3), and even
Territorial use communal rights (GS6.1.4.3). However, we identified
important differences in terms of the actors, resource systems, and
resource units. Although all communities show a high level of leadership
(A5), the level of trust (A6.1) is lower in Punta Herrero. In general,
the presence and success of conservation initiatives depends on the
incentives of local communities to maintain a healthy status of the
resources they depend upon \citep{jupiter_2017}. The enabling conditions
for conservation seem to be strongly present in all communities. Due to
the clarity of access rights and isolation, the benefits of conservation
directly benefit the members of the fishing cooperative. These
conditions have favored the development of efficient community-based
enforcement systems.

@@

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{\label{fig:unnamed-chunk-7}\label{fig:lobsters}Effect sizes for
lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
circles) and Maria Elena (ME; blue triangles). Points indicate the
effect size and standard errors. Years have been centered to year of
implementation.}
\end{figure}

@@

Some works evaluate marine reserves by performing inside-outside
\citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD} or
before-after comparisons \citep{betti_2017-lq}. The first approach does
not address temporal variability, and the second can not distinguish
between the temporal trends in a reserve and the entire system
\citep{depalma_2018}. Our approach to evaluate the temporal and spatial
changes provides a more robust measure of reserve effectiveness. For
example, we capture previously described patterns like the rapid
increase observed for changes in lobster densities for Isla Natividad on
the sixth year (\emph{i.e.} 2012; Fig. \ref{fig:indicators}A) occur a
year after the hypoxia events described by \citet{micheli_2012-EU},
which caused mass mortality of sedentary organisms such as abalone and
sea urchins, but not lobster and finfish. Yet, our empirical approach
assumes control sites are a plausible counterfactual for treated sites.
This implies that treated sites would have followed the same trend as
control sites, had the reserves not been implemented. Nonetheless,
temporal trends for each site don't show any significant increases,
supporting our findings of lack of change in the indicators used.

@@

Another key condition for effectiveness is reserve size
\citep{edgar_2014-UO}, and the lack of effectiveness can perhaps be
attributed to poor ecological coherence in reserve design (\emph{sensu}
\citet{rees_2018}. Previous research has shown that reserves in Isla
Natividad yield fishery benefits for the abalone fishery
\citep{rossetto_2015-V0}. Abalone are less mobile than lobsters, and
perhaps the reserves provide enough protection to these sedentary
invertebrates, but not lobsters. Design principles developed by
\citet{green_2017} for marine reserves in the Caribbean state that
reserves ``should be more than twice the size of the home range of
adults and juveniles'', and suggest that reserves seeking to protect
spiny lobsters should have at least 14 km across. Furthermore, may favor
implementation of reserves that pose low fishing costs due to their
small size or location. Our analysis of economic data supports this, as
neither landings nor revenues showed the expected short-term costs
associated to the first years of reserve implementation
\citep{ovando_2016-Wg}.

@@

While the evaluated reserves have failed to provide fishery benefits up
to now, there are a number of additional ecological, fisheries, and
social benefits. Marine reserves provide protection to a wider range of
species and vulnerable habitat, like coral reefs. These sites can serve
as an insurance against uncertainty and errors in fisheries management,
as well as environmental shocks
\citep{hilborn_2004,hilborn_2006,micheli_2012-EU} (\textbf{,Aalto et al,
in press}). Self-regulation of fishing effort (\emph{i.e.} reduction in
harvest) can serve as a way to compensate for future declines associated
to environmental variation \citep{finkbeiner_2018}. Embarking in a
marine conservation project can bring the community together, which
promotes social cohesion and builds social capital \citep{fulton_2019}.
Furthermore, showing commitment to marine conservation and sustainable
fishing practices allows fishers to have greater bargaining power and
leverage over fisheries management \citep{prezramrez_2012}.

@@

JCVD recevied funding from CONACyT (Beca de Posgrados en el extranjero,
CVU 669403) and the Latin American Fisheries Fellowship Program. AS,
AHV, SF and JT received funding from Marisla Foundation, Packard
Foundation, Walton Family Foundation, Summit Foundation, and Oak
Foundation.

@@
